---
title: "hw2reporoducibleresearch"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
This assignment makes use of data from a personal activity monitoring device. This device collects data at 5 minute intervals through out the day. The data consists of two months of data from an anonymous individual collected during the months of October and November, 2012 and include the number of steps taken in 5 minute intervals each day.

The data for this assignment can be downloaded from the course web site:

Dataset: Activity monitoring data [52K]
The variables included in this dataset are:

steps: Number of steps taking in a 5-minute interval (missing values are coded as NA)
date: The date on which the measurement was taken in YYYY-MM-DD format
interval: Identifier for the 5-minute interval in which measurement was taken
The dataset is stored in a comma-separated-value (CSV) file and there are a total of 17,568 observations in this dataset.

# Loading and preprocessing data 

``` {r echo= TRUE}
## Set the working directory
setwd("C://Users//Sudha Kumar//Documents//Coursera//reproducibleresearch//")

## Loading libraries
library(dplyr)
library(ggplot2)

## read the data in the dataframe
df1<-read.csv("activity.csv", header= TRUE, sep=",")

```
# What is mean total number of steps taken per day

```{r echo= TRUE}
tosum<-df1%>% group_by(date)%>% summarize(tosum=sum(steps,na.rm= TRUE))


# Make a histogram of the total number of steps taken each day

hist(tosum$tosum,main="Total number of steps",col="red",xlab="Total Number of steps per day")

# Calculate and report the mean and median of the total number of steps taken per day

stepsmean<-mean(tosum$tosum)
# The mean of total number of steps is 9354

stepsmedian<-median(tosum$tosum)

# The mean of total number of steps is 10395
```
# What is the average daily activity pattern?

```{r echo=TRUE}
# Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

#Time series plot of average steps per day
stepsperinterval<-aggregate(steps~interval, data=df1,FUN =mean)
ggplot(stepsperinterval,aes(interval,steps))+geom_line()
#max interval
max_interval <- stepsperinterval[which.max(stepsperinterval$steps),1]

```
# Imputing missing values

```{r echo= TRUE}
#if the data is NA change to one, the reason is there is a margin of error and there i s always one step that was taken
imputed_data <- transform(df1,steps=ifelse(is.na(df1$steps),1,df1$steps))

imputedsum<-imputed_data%>% group_by(date)%>% summarize(imputedtosum=sum(steps))

hist(imputedsum$imputedtosum,main="Total number of steps",col="red",xlab="Total Number of steps per day")

#imputed mean

imputedstepsmean<-mean(imputedsum$imputedtosum)
# Imputed mean is 9392

imputedstepsmedian<-median(imputedsum$imputedtosum)
# imputed media is 10395

```
# Are there differences in activity patterns between weekdays and weekends?

```{r echo=FALSE}
weekdays <- c("Monday", "Tuesday", "Wednesday", "Thursday", 
              "Friday")
imputed_data$dow = as.factor(ifelse(is.element(weekdays(as.Date(imputed_data$date)),weekdays), "Weekday", "Weekend"))

steps_by_interval_i <- aggregate(steps ~ interval + dow, imputed_data, mean)

library(lattice)

xyplot(steps_by_interval_i$steps ~ steps_by_interval_i$interval|steps_by_interval_i$dow, main="Average Steps per Day by Interval",xlab="Interval", ylab="Steps",layout=c(1,2), type="l")

```